---
title: "ScDblFinder filtering with recoverDoublets - `r params$sample_name`"
author: 'Antonia Chroni <chronia@chop.edu> for D3B'
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: TRUE
    toc_float: TRUE
params:
  data_path: './' # path to the directory with the `seurat_obj_raw.rds` file from the `seurat_alignment_qc.Rmd` script
  results_dir: './' # path to dir to save all output files
  sample_name: "pbmc_1k_v3"   # name of the sample in the library
  log_file: NULL
---

# Information about this notebook
This script used the [ScDblFinder](https://bioconductor.org/packages/devel/bioc/vignettes/scDblFinder/inst/doc/scDblFinder.html) method for identifying doublets in single-cell data. 

# Usage
To run the Rscript from the command line sequentially, use:

```
Rscript -e "rmarkdown::render('scDblFinder.Rmd', clean = TRUE,
      params=list(results_dir = '.', 
                  data_path = '/path/to/seurat_alignment_qc/output'),
                  sample_name = 'pbmc_1k_v3')"
```

# Set up

```{r load-library, echo=TRUE}
library(scDblFinder)
library(Seurat)
library(scater)
library(future)

# # evaluate Seurat R expressions asynchronously when possible (such as ScaleData) using future package
plan("multisession", workers = 4)
options(future.globals.maxSize = 64000 * 1024^2) # for 64 Gb RAM
```

# Directories and paths to file Inputs/Outputs

```{r set-dir-and-file-names, echo=TRUE}
attach(params)

# File path to the directory with the outs file from the `seurat_alignment_qc.Rmd` script
data_dir <- c(dir(path = data_path,  pattern = "seurat_obj_raw.rds", full.names = TRUE, recursive = TRUE))

#create results_dir
results_dir <-
  file.path(results_dir)
if (!dir.exists(results_dir)) {
  dir.create(results_dir)
}


# File path to doublets directory
doublets_dir <- file.path(paste0(results_dir,"/scDblFinder-", sample_name, "/"))
if (!dir.exists(doublets_dir)) {
  dir.create(doublets_dir)
}

knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.path = file.path(results_dir))

```


# Read raw seurat_objs 
We will load the raw seurat object as generated in the `seurat_alignment_qc.Rmd` script.
We will run dimensionality reductions and convert to single cell experiments.


```{r process-seurat-obj-convert-to-sce, echo=TRUE}
seurat_obj <- readRDS(data_dir)
DefaultAssay(seurat_obj) <- "RNA"

# Make this reproducible as UMAP algorithm is not deterministic
set.seed(1234)

# Normalize, find variable features and scale data
seurat_obj <- seurat_obj_raw %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData() 

# Run dim_reductions
seurat_obj <- RunPCA(seurat_obj, features = VariableFeatures(object = seurat_obj), npcs=30) %>%
                    RunUMAP(dims = 1:20, n.neighbors=20L, reduction.name = "umap", reduction.key = "UMAP_", verbose = FALSE)

seurat_obj <- as.SingleCellExperiment(seurat_obj)

# Save sce
saveRDS(seurat_obj, file = paste0(results_dir, "/", "seurat_obj_sce.rds"))
```

# Run scDblFinder 

`scDblFinder function` is used to identify cells that are doublets based on cluster-based generation of artificial doublets. 
We will save the cells that need to be filtered based on scDblFinder analysis in a separate table.


```{r run-recoverDoublets, echo=TRUE}
# Run scDblFinder
bp <- BiocParallel::MulticoreParam(4, RNGseed=1234)
BiocParallel::bpstart(bp)
set.seed(1234)
seurat_obj  <- scDblFinder(seurat_obj, clusters = TRUE, BPPARAM = bp)
BiocParallel::bpstop(bp)

# Estimate singlets and doublets in the library
singlets_doublets_number <- print(table(seurat_obj$scDblFinder.class))

# Estimate pct of doublets in the library
doublets_pct_library <- table(seurat_obj$scDblFinder.class)[2]/(table(seurat_obj$scDblFinder.class)[1]+table(seurat_obj$scDblFinder.class)[2])*100

doublets <- seurat_obj@colData[,str_detect(colnames(seurat_obj@colData), "scDblFinder")] %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column("cell")

# Save table with all predictions
write.csv(doublets, paste0(doublets_dir, "doublets_", sample_name, ".csv"))
    
# Cells that we need to filter from the final object
doublets.to.filter <- subset(doublets, doublets$scDblFinder.class == "doublet")

doublets_number <- print(table(seurat_obj$scDblFinder.class))

# Save table with cells to be filtered
write.csv(doublets.to.filter, paste0(doublets_dir, "doublets_to_filter_", sample_name,".csv"))

# Save seurat_obj with doublets predictions in the metadata
saveRDS(seurat_obj, paste0(doublets_dir, "sce_obj_", sample_name,".rds") )
```

There are `r singlets_doublets_number` singlets and doublets, respectively. 
Doublets represent `r doublets_pct_library` % of cells in the whole library.


# Plot the Doublets predictions on the UMAP

```{r plot-Doublets, echo=TRUE}
# Let's plot the Doublets predictions on the UMAP
print("Doublets predictions")
print(gridExtra::grid.arrange(plotUMAP(seurat_obj, colour_by="scDblFinder.cluster", point_size = 0.1) + ggtitle("Clusters"),
                              plotUMAP(seurat_obj, colour_by="scDblFinder.score", point_size = 0.1) + ggtitle("Doublet score"),
                              plotUMAP(seurat_obj, colour_by="scDblFinder.class", point_size = 0.1) + ggtitle("Predicted doublets"),
                              ncol=2))

pdf(paste0(doublets_dir, "Doublets_predictions.pdf"), onefile=FALSE, useDingbats=F, height=12, width=12)
print(gridExtra::grid.arrange(plotUMAP(seurat_obj, colour_by="scDblFinder.cluster", point_size = 0.1) + ggtitle("scDblFinder clusters"),
                              plotUMAP(seurat_obj, colour_by="scDblFinder.score", point_size = 0.1) + ggtitle("Doublets score"),
                              plotUMAP(seurat_obj, colour_by="scDblFinder.class", point_size = 0.1) + ggtitle("Doublets predictions"),
                              ncol=2))
dev.off()
```


```{r}
sessionInfo()
```

